The CrypTool Script: \\ Cryptography, Mathematics and More \\ \vspace{30pt} {\large Background reading\\%*[-8pt] for the free e-learning program CrypTool\\ (10th edition -- distributed with CrypTool version 1.4.30)} \vspace{60pt}

### 2013-06-20 script/de/chapters/numbertheory.tex
CT Script: Minor changes.

\begin{satz}\label{thm-zth-div}
$a \equiv b$ (mod $m$) gilt genau dann,  wenn die Differenz $(a - b)$ durch $m$ teilbar ist, also wenn 
ein $q\in \mathbf{Z}$ existiert mit $ (a-b)=q*m.$
\end{satz}
Diese beiden Aussagen sind also äquivalent.\footnote{%
Die obige Äquivalenz gilt nur für die Differenz $(a - b)$, nicht für die
Summe $(a + b)$!

@@

\noindent Für Summen gilt die Aussage von Satz~\ref{thm-zth-div} nicht einmal
in eine Richtung. Richtig ist sie bei Summen nur für den Rest $0$ und nur in
der folgenden Richtung: Teilt ein Teiler beide Summanden ohne Rest, teilt er
auch die Summe ohne Rest.
}

>>>>

\begin{satz}\label{thm-zth-div}
$a \equiv b$ (mod $m$) gilt genau dann,  wenn die Differenz $(a - b)$ durch $m$
teilbar ist, also wenn ein $q\in \mathbf{Z}$ existiert mit $ (a-b)=q*m.$\footnote{%
Die obige Äquivalenz gilt nur für die Differenz $(a - b)$, nicht für die
Summe $(a + b)$!

@@

\noindent Für Summen gilt die Aussage von Satz~\ref{thm-zth-div} nicht einmal
in eine Richtung. Richtig ist sie bei Summen nur für den Rest $0$ und nur in
der folgenden Richtung: Teilt ein Teiler beide Summanden ohne Rest, teilt er
auch die Summe ohne Rest.
}
\end{satz}

\noindent In anderen Worten:~~~
$ a \equiv b \bmod{m}   ~~ \Longleftrightarrow ~~
  m | (a-b)             ~~ \Longleftrightarrow ~~
  (a-b) \equiv 0 \bmod{m} $

### 2013-06-20 script/en/chapters/numbertheory.tex
CT Script: Minor changes.

\begin{theorem}\label{thm-zth-div}
$a \equiv b$ (mod $m$) if and only if, the difference $(a - b)$ is divisible by $m$, i.e. if $q\in 
\mathbb{Z}$ exists with $ (a-b)=q*m.$
\end{theorem}
These two statements are therefore equivalent.\footnote{%
The above equivalence does apply only to the difference $(a - b)$, not to the
sum $(a + b)$!

@@

\noindent The statement in theorem~\ref{thm-zth-div} does not even apply to
sums in one direction.
It is correct for sums only if the remainder is 0 and only in the following direction:
If a divisor divides both summands with no remainder, it also divides the sum with no remainder.
}

@@

The paper~\cite{nt:Heninger2012} explains well \textit{how} the algorithm works, but not as well \textit{why}. The product $P$ of all moduli is a very big number, even compared to a single modul. Without the simplifications from the remainder tree you would go the following way: Calculate $gcd_{i} = gcd( P / m_{i},  m_{i}) $ for all i. Compare each $gcd_{i} \ne 1$ with all other $gcd_{j} \ne 1$ with  $j \ne i$. If two gcd's are the same, then their moduli share a factor.

@@

Within the remainder tree you get -- at the end -- $(P \bmod ({m_{i}^{2}}) ) / m_{i}$ for all $i$.\footnote{%
It would not make sense to calculate $(P \bmod m_{i} ) / m_{i}$
as $(P \bmod m_{i} )$ is always $= 0$.

@@

$$gcd((P \bmod {m_{i}^{2}})/ m_{i}, m_{i})~~~  \overset{!}{=}  ~~~gcd( P / m_{i}, m_{i})$$
$ \Longleftrightarrow $~~~\footnote{According to Euklid's algorithm the following identities are true:
\mbox{}\\
$gcd((P \bmod {m_{i}^{2}})/ m_{i}, m_{i})=gcd((P \bmod {m_{i}^{2}})/ m_{i} \bmod{m_{i}}, m_{i})$ \mbox{}\\
$gcd( P / m_{i}, m_{i})=gcd( P / m_{i} \bmod{m_{i}}, m_{i})$\mbox{}
}

@@

$$(P \bmod{m_{i}^{2}})/ m_{i} - P / m_{i} \equiv 0   \bmod{m_{i}}~~~ \Leftrightarrow ~~~  m_{i} ~~ | ~~ ((P \bmod{m_{i}^{2}})/ m_{i} - P / m_{i})$$~~~\footnote{%
According to the definition of modulus and division it is:\\
$a \bmod{b} ~~ \Longleftrightarrow ~~ a - b * \lfloor a/b \rfloor$\\
So $P \bmod{m_{i}^{2}}$ can be written as $ P-m_{i}^{2} \lfloor P/m_{i}^{2} \rfloor$.
}

>>>>

\begin{theorem}\label{thm-zth-div}
$a \equiv b$ (mod $m$) if and only if, the difference $(a - b)$ is divisible by
$m$, i.e. if $q\in \mathbb{Z}$ exists with $ (a-b)=q*m.$\footnote{%
The above equivalence does apply only to the difference $(a - b)$, not to the
sum $(a + b)$!

@@

\noindent The statement in theorem~\ref{thm-zth-div} does not even apply to
sums in one direction.
It is correct for sums only if the remainder is 0 and only in the following direction:
If a divisor divides both summands with no remainder, it also divides the sum with no remainder.
}
\end{theorem}

\noindent In other words:~~~
$ a \equiv b \bmod{m}   ~~ \Longleftrightarrow ~~
  m | (a-b)             ~~ \Longleftrightarrow ~~
  (a-b) \equiv 0 \bmod{m} $

@@

The paper~\cite{nt:Heninger2012} explains well \textit{how} the algorithm works, but not as well \textit{why}. The product $P$ of all moduli is a very big number, even compared to a single modul. Without the simplifications from the remainder tree you would go the following way: Calculate $gcd_{i} = gcd( P / m_{i},  m_{i}) $ for all i. Compare each $gcd_{i} \ne 1$ with all other $gcd_{j} \ne 1$ with  $j \ne i$. If two gcd's are the same, then their moduli share a factor.\footnote{%
A prerequisite for getting only prime factors, is that duplicate moduli are removed
before setting up the trees.
}

@@

Within the remainder tree you get -- at the end -- $(P \bmod ({m_{i}^{2}}) ) / m_{i}$ for all $i$.\footnote{%
It would not make sense to calculate $(P \bmod m_{i} ) / m_{i}$, as 
$m_{i} | P$, so $P/m_{i}$ is always a whole number, which means  
$(P \bmod m_{i} )$ is always $= 0$.

@@

$$gcd((P \bmod {m_{i}^{2}})/ m_{i}, m_{i})~~~  \overset{!}{=}  ~~~gcd( P / m_{i}, m_{i})$$
$ \Longleftrightarrow $~~~\footnote{According to Euklid's algorithm (first iteration) the following identity is true:\\
$gcd(a, b) = gcd(b, a \bmod b)$ if $ b \neq 0 $\\
This holds as per definition it is: $gcd(a, 0) = a$\\
Applied to our problem this means:
\mbox{}\\
$gcd((P \bmod {m_{i}^{2}})/ m_{i}, m_{i})=gcd((P \bmod {m_{i}^{2}})/ m_{i} \bmod{m_{i}}, m_{i})$ \mbox{}\\
$gcd( P / m_{i}, m_{i})=gcd( P / m_{i} \bmod{m_{i}}, m_{i})$\mbox{}
}

@@

$$(P \bmod{m_{i}^{2}})/ m_{i} - P / m_{i} \equiv 0   \bmod{m_{i}}~~~ \Leftrightarrow ~~~  m_{i} ~~ | ~~ ((P \bmod{m_{i}^{2}})/ m_{i} - P / m_{i})$$~~~\footnote{%
According to the definition of modulus (see theorem \ref{thm-zth-div}) and
division it is:
$a \bmod{b} ~~ \Longleftrightarrow ~~ a - b * \lfloor a/b \rfloor$\\
So $P \bmod{m_{i}^{2}}$ can be written as $ P-m_{i}^{2} \lfloor P/m_{i}^{2} \rfloor$.
}
